Bayesian Sparsification of Deep $\mathbb{C}$-valued Networks

### 2020-06-18 paper/paper.tex
icml2020 header

% The Local Reparameterization Trick reduces the variance of SGVB
% even further by sampling independent weight matrices for
% each data-point inside mini-batch. So we need both original SGVB
% (per sample randomness) and local reparameterization (that is an
% efficient and less noisy equivalent to per-sample weight sampling)
% SGVB is specifically for VAE (local latent variables per data-point)
% what Kingma propose is SGVB for global shared latent variables aka
% WEIGHTS!!
In this study we use the SGVB estimator \eqref{eq:sgvb_estimator} with $L=1$ and the
\emph{local reparameterization trick} proposed by \citet{kingma_variational_2015}. They
argued that this gradient estimator can be made more statistically and computationally
efficient, if the structure of the model permits translating global stochasticity of
$\omega$ down to local intermediate states of computation.
%
The class of models that allow this include non-recurrent computational graphs,
exemplified by neural networks with parameters $\omega \sim q_\theta$. In their case,
\eqref{eq:sgvb_estimator} would require that the \emph{entire} set of network's parameters
$\omega$ be independently drawn for each element in the mini-batch. Since the parameters
in a network naturally split into subsets with non-overlapping layer-wise effects, it is
standard to assume that the approximation $
  q_\theta(\omega)
$ is factorized over layers. Furthermore, if $
  % \omega =
  W \in \mathbb{R}^{n\times m}
$ in a linear layer $
  y = b + W^\top x
$ with $
  q_\theta(W)
    = \prod_{ij}
      \mathcal{N}(w_{ij} \vert\, \mu_{ij}, \sigma_{ij}^2)
$, then by virtue of $y$ being a linear transformation of $W$, we get
\begin{equation}  \label{eq:r-gauss-trick}
  q(y) = \prod_i \mathcal{N}\Bigl(
        y_i \big\vert\,
        b_i + \sum_j \mu_{ij} x_j,
        \sum_j \sigma^2_{ij} x_j^2
    \Bigr)
  \,.
\end{equation}
This yields outputs equivalent in distribution to sampling $W$ for each element in
the mini-batch, which produces the SGVB estimator with smaller variance, as demonstrated
by \citet{kingma_variational_2015}.
% Similar idea to \eqref{eq:r-gauss-trick} was proposed by \citet{wang_fast_2013} to speed
% up Gaussian dropout.

@@

% $\cplx$-Gaussians are closed under affine transformations
$\cplx$-Gaussianity is preserved under linear transformations, i.e. for $
  A \in \cplx^{n \times m}
$ and $b \in \cplx^{n}$
\begin{equation}  \label{eq:cn-affine}
  b + A z \sim \mathcal{C}\mathcal{N}_n\bigl(
      b + A\mu, A \Gamma \conj{A^\top}, A C A^\top
    \bigr)
  \,.
\end{equation}
Therefore, if we have a $\cplx^{n\times m}$ matrix $W$ with \emph{independent}
$\cplx$-Gaussian entries, i.e.
\begin{equation}  \label{eq:c-gauss-vi-general}
  q(W)
    = \prod_{ij} \mathcal{CN}(
      % W_{ij} \mid
      \mu_{ij}, \Sigma_{ij}, \Sigma_{ij} \xi_{ij}
    )
  \,,
\end{equation}
with $
  \mu, \xi \in \cplx^{n\times m}
$, $\Sigma \in [0, +\infty)^{n\times m}$ and $
  \lvert \xi_{ij} \rvert \leq 1
$, then for $x \in \cplx^m$ and $b \in \cplx^n$ each component $y_i$ of $y = b + W x$
is independent univariate $\cplx$-Gaussian
\begin{equation}  \label{eq:cplx-gauss-trick}
  y_i
    \sim \mathcal{C}\mathcal{N}
      \Bigl(
        b_i + \sum_{j=1}^m \mu_{ij} x_j,
        \, \sum_{j=1}^m \Sigma_{ij} \lvert x_j \rvert^2,
        \, \sum_{j=1}^m \Sigma_{ij} x_j^2 \xi_{ij}
      \Bigr)
    \,.
\end{equation}
This is the $\cplx$-Gaussian version of the local reparameterization trick \eqref{eq:r-gauss-trick}.
It requires three matrix-vector operations: $\cplx$-valued $b + \mu x$ and $C x^2$,
and $\real$-valued $
  \Sigma \lvert x \rvert^2
$, where $C_{ij} = \Sigma_{ij} \xi_{ij}$ and the complex modulus and square are applied
elementwise. Equation \eqref{eq:cplx-gauss-trick} can be applied to any layer, the output
of which depends linearly on its parameters, such as convolutional, affine, and bilinear
transformations ($
  (x, z) \mapsto b_j + x^\top W^{(j)} z
$). Similar to the $\real$ case, $\cplx$ convolutions draw independent realizations of $W$
for each spatial patch in the input \citep{molchanov_variational_2017}. This provides
faster computations and better statistical efficiency of the SGVB gradient estimator
by eliminating correlation from overlapping patches \citep{kingma_variational_2015} and
allowing \eqref{eq:cplx-gauss-trick} to efficiently leverage $\cplx$ convolutions of
the relation and variance kernels with elementwise complex squares $x^2$ and amplitudes
$\lvert x \rvert^2$.
% Weight sharing is essentially pushed to shared variational parameters $\theta$ in the approximation $q_\theta$.
% A convolution is a matrix product of $W$, unrolled into a toeplitz matrix, and im-to-col
% flattened intput $x$

>>>>

% The Local Reparameterization Trick reduces the variance of SGVB
% even further by sampling independent weight matrices for
% each data-point inside mini-batch. So we need both original SGVB
% (per sample randomness) and local reparameterization (that is an
% efficient and less noisy equivalent to per-sample weight sampling)
% SGVB is specifically for VAE (local latent variables per data-point)
% what Kingma propose is SGVB for global shared latent variables aka
% WEIGHTS!!
In this study we use the SGVB estimator \eqref{eq:sgvb_estimator} with $L=1$ and the
\emph{local reparameterization trick} proposed by \citet{kingma_variational_2015}. They
argued that this gradient estimator can be made more statistically and computationally
efficient, if the structure of the model permits translating global stochasticity of
$\omega$ down to local intermediate states of computation.
%
The class of models that allow this include non-recurrent computational graphs,
exemplified by neural networks with parameters $\omega \sim q_\theta$. In their case,
\eqref{eq:sgvb_estimator} would require that the \emph{entire} set of network's parameters
$\omega$ be independently drawn for each element in the mini-batch. Since that the
parameters in a network naturally split into subsets with non-overlapping layer-wise
effects, it is standard to assume that the approximation $
  q_\theta(\omega)
$ is factorized over layers. Furthermore, if $
  % \omega =
  W \in \mathbb{R}^{n\times m}
$ in a linear layer $
  y = b + W^\top x
$ with $
  q_\theta(W)
    = \prod_{ij}
      \mathcal{N}(w_{ij} \vert\, \mu_{ij}, \sigma_{ij}^2)
$, then by virtue of $y$ being a linear transformation of $W$, we get
\begin{equation}  \label{eq:r-gauss-trick}
  q(y) = \prod_i \mathcal{N}\Bigl(
        y_i \big\vert\,
        b_i + \sum_j \mu_{ij} x_j,
        \sum_j \sigma^2_{ij} x_j^2
    \Bigr)
  \,.
\end{equation}
This yields outputs equivalent in distribution to sampling $W$ for each element in
the mini-batch, which produces the SGVB estimator with smaller variance, as demonstrated
by \citet{kingma_variational_2015}.
% Similar idea to \eqref{eq:r-gauss-trick} was proposed by \citet{wang_fast_2013} to speed
% up Gaussian dropout.

@@

% $\cplx$-Gaussians are closed under affine transformations
$\cplx$-Gaussianity is preserved under linear transformations, i.e. for $
  A \in \cplx^{n \times m}
$ and $b \in \cplx^{n}$
\begin{equation}  \label{eq:cn-affine}
  b + A z \sim \mathcal{C}\mathcal{N}_n\bigl(
      b + A\mu, A \Gamma \conj{A^\top}, A C A^\top
    \bigr)
  \,.
\end{equation}
Therefore, if we have a $\cplx^{n\times m}$ matrix $W$ with \emph{independent}
$\cplx$-Gaussian entries, i.e.
\begin{equation}  \label{eq:c-gauss-vi-general}
  q(W)
    = \prod_{ij} \mathcal{CN}(
      % W_{ij} \mid
      \mu_{ij}, \Sigma_{ij}, \Sigma_{ij} \xi_{ij}
    )
  \,,
\end{equation}
with $
  \mu, \xi \in \cplx^{n\times m}
$, $\Sigma \in [0, +\infty)^{n\times m}$ and $
  \lvert \xi_{ij} \rvert \leq 1
$, then for $x \in \cplx^m$ and $b \in \cplx^n$ each component $y_i$ of $y = b + W x$
is independent univariate $\cplx$-Gaussian
\begin{equation}  \label{eq:cplx-gauss-trick}
  y_i
    \sim \mathcal{C}\mathcal{N}
      \Bigl(
        b_i + \sum_{j=1}^m \mu_{ij} x_j,
        \, \sum_{j=1}^m \Sigma_{ij} \lvert x_j \rvert^2,
        \, \sum_{j=1}^m \Sigma_{ij} x_j^2 \xi_{ij}
      \Bigr)
    \,.
\end{equation}
This is the $\cplx$-Gaussian version of the local reparameterization trick \eqref{eq:r-gauss-trick}.
It requires three matrix-vector operations: $\cplx$-valued $b + \mu x$ and $C x^2$,
and $\real$-valued $
  \Sigma \lvert x \rvert^2
$, where $C_{ij} = \Sigma_{ij} \xi_{ij}$ and the complex modulus and square are applied
elementwise. \eqref{sub:c_gauss_and_local_rep} can be applied to any layer, the output
of which depends linearly on its parameters, such as convolutional, affine, and bilinear
transformations ($
  (x, z) \mapsto b_j + x^\top W^{(j)} z
$). Similar to the $\real$ case, $\cplx$ convolutions draw independent realizations of $W$
for each spatial patch in the input \citep{molchanov_variational_2017}. This provides
faster computations and better statistical efficiency of the SGVB gradient estimator
by eliminating correlation from overlapping patches \citep{kingma_variational_2015} and
allowing \eqref{eq:cplx-gauss-trick} to efficiently leverage $\cplx$ convolutions of
the relation and variance kernels with elementwise complex squares $x^2$ and amplitudes
$\lvert x \rvert^2$.
% Weight sharing is essentially pushed to shared variational parameters $\theta$ in the approximation $q_\theta$.
% A convolution is a matrix product of $W$, unrolled into a toeplitz matrix, and im-to-col
% flattened intput $x$
